Quantifying and controlling biases in dark matter halo concentration estimates

### 2016-07-14 paper/ms.tex
notes on the approximated covariance matrix

We divide the enclosed mass mass $M_i$ and the radii $r_i$ by their
virial values to obtain the dimensionless variables $m_i$ and $x_i$.
Once the mass profile is expressed in dimensionless variables the
concentration is the only free parameter.  
We then use an Affine Invariant Markov Chain Monte Carlo implemented
in the python module {\texttt{emcee}} \citep{emcee} to sample the
likelihood function distribution defined by ${\cal L}(c)\propto
\exp(-\chi^2(c)/2)$ where the $\chi^2(c)$ is written as

\begin{equation}
\chi^2(c)= \frac{1}{N}\sum_{i=1}^{N}[\log m_i - \log m(< x_i;c)]^2,
\end{equation}
%
where $m(<x_i;c)$ corresponds to the values in Eq.(\ref{eq:profile})
at $x=x_i$ for a given value of the concentration parameter $c$ and
the $i$ index sums over all the particles in the numerical profile.
From the $\chi^2$ distribution we find the optimal value of the
concentration and its associated uncertainty.

>>>>

We divide the enclosed mass mass $M_i$ and the radii $r_i$ by their
virial values to obtain the dimensionless variables $m_i$ and $x_i$.
Once the mass profile is expressed in dimensionless variables the
concentration is the only free parameter.

Using the bootstrapping data we find that at a given normalized
radius, $x$, the logarithm of the normalized integrated mass, $m$, follows a 
distribution that can approximated by a gaussian with variance
\begin{equation}
\sigma_x^2 = \frac{1-x}{x}\frac{1}{6N}.
\label{eq:sigma}
\end{equation}

If the integrated mass values at different radii were independent from
each other, this could be used as the basis to develop a simple
bayesian likelihood to be sampled with Markov Chain Monte Carlo techniques where
the likelihood could be then defined to be ${\cal L}(c)\propto
\exp(-\chi^2(c)/2)$ with a $\chi^2(c)$ written as

\begin{equation}
\chi^2(c)= \sum_{i=2}^{N-1}\frac{[\log m_i - \log m(< x_i;c)]^2}{\sigma_i^2},
\end{equation}
%
where $\sigma_i^2=\sigma_x^2$, $m(<x_i;c)$ corresponds to the
values in Eq.(\ref{eq:profile}) at $x=x_i$ for a given value of the
concentration parameter $c$, and the $i$ index sums over all the
particles in the numerical profile. 
In this computation the particles $i=1$ and $i=N$ are discarded to
avoid divergent terms in the sum.

However, tests on the bootstrapping data show that the covariance
matrix is not diagonal and using $\sigma_i^2=\sigma_x^2$ would grossly
overestimate $\chi^2(c)$ providing too optimistic error bars.

Given the high expense of computing a full covariance matrix for
each desired number of particles $N$, we have calibrated an effective
$\sigma_{\rm eff}^2$ such that it keeps the dependence on $x$ that we have
discovered for the diagonal elements and also gives similar curves of
$\chi^{2}(c)$-$c$ around the minimum as the full covariance matrix
computation.

We found that this effective $\sigma_{\rm eff}^2$ can be approximated as

\begin{equation}
\sigma_{\rm eff}^2 = \frac{1-x}{x}\frac{N^{1.15}}{4.5\times 10^3}.
\end{equation}

We then use $\sigma_i^2=\sigma_{\rm eff}^2$ to feed an Affine
Invariant Markov Chain Monte Carlo implemented in the python module
{\texttt{emcee}} \citep{emcee} to sample the likelihood function
distribution.
From the $\chi^2$ distribution we find the optimal value of the
concentration and its associated uncertainty.
